---
title: "One Dose Response Curve"
output: html_One_Dose_Response_Curve
---

### In the first  tutorial, we will learn how to fit a four parameter log-logistic using the data from the pig ATPase and estimate the model parameters (Figure 3a) ### 

Step 1. Reading a data file into R. We will first install the packages tidyverse and drc using the install.packages() function. The library(drc) and the library(tidyverse) functions load and activate the packages drc and tidyverse in R
```{r}
install.packages("tidyverse")
install.packages("drc")
library(tidyverse)
library(drc)
```

Step 2. Assign the name DoseResponse_Data_Pig to your data file and subset with the filter() function from the dplyr() package of tidyverse the rows with observations from the pig ATPase. Use the head() function to print the first 6 rows of the dataframe. This is useful to check if the data have been read correctly.
```{r}
DoseResponse_Data_Pig<- read.csv("Figure3a.csv") %>% 
  dplyr::filter(Species_ATPase == "Pig ATPase")

head(DoseResponse_Data_Pig)
```
Step 3. Plot the activity of the pig ATPase of the pig as a function of the ouabain dose both in the arithmetic and logarithmic scales. Notice that when we convert the X-axis values (Dose) from arithmetic to logarithmic scale, the plotted dose-response curve will be changed from hyperbolic to sigmoid ("S" shape, stretched). Usually, between 25% to 75 % of the Maximum response, the relation between doses and responses will be linear in the logarithmic scale. A better understanding and interpretation can be drawn from this linear area. 
```{r}
op <- par(mfrow = c(1, 2), mar=c(3.2,3.2,2,.5), mgp=c(2,.7,0)) #make two plots in two columns 
plot(Activity ~ Dose, data = DoseResponse_Data_Pig, main="Original Dose Scale")
plot(Activity ~ log10(Dose), data = DoseResponse_Data_Pig, main="Logarithmic Dose Scale")
```

Step 4. Fit a log logistic dose-response curve to the data of the ping ATPase. The drm() function is the key function in the drc package for fitting dose-response curves. Activity is the response variable (y-axis) and Dose is the dose (x-axis).Species ATPase is the classification variable. data= DoseResponse_Data_Pig identifies the name of the data file. fct=LL.4() is the argument for the log-logistic curve with four parameters.  Here, we fixed two of the parameters, the lower and upper limits, at 0 and 100, respectively.Notice that the drm() function does not produce an output. All the information on the model fit is stored in the object under our defined name DoseResponse_Data_Pig.model. This will be read by further commands.
```{r}
DoseResponse_Data_Pig.model <- drm(Activity ~ Dose, 
                                   data = DoseResponse_Data_Pig, 
                                   fct = LL.4(fixed = c(NA, 0,100, NA)))

```

Step 5. The summary() function provides a summary of the parameter estimates for the DoseResponse_Data_Pig.model. In this case, we have constrained the lower and upper limits to 0 and 100, respectively. The parameters that are estimated are the slope (b) and the effective dose (e) and their standard errors. The p-values tell us if the parameters are different from zero.
```{r}
summary(DoseResponse_Data_Pig.model)
```

Step 6. The ED() function calculates the estimated effective dose based on the fitted model.
```{r}
ED(DoseResponse_Data_Pig.model, c(50))
```

Step 7. We use the plot() function from base R to obtain visuals of the fitted curve. The argument "type = 'all'" displays all data points on the plot because the default is to plot the mean. The argument col() provides the color of the data. The arguments xlab() and ylab() provide the labels on the axes. The argument main() provides the title of the plot. Notice that by default plot() uses a logarithmic dose axis. This can be switched off with the argument log = "".
```{r}
plot(DoseResponse_Data_Pig.model,
     type = "all",
     lwd = 2,
     lty=3 ,
     col = "black",
     main="Pig ATPase",
     xlab = "Dose (M)", 
     ylab = "Na+/K+-ATPase activity \n (% of non-inhibited control)")
```
Step 8. We can also use the ggplot() package from tidyverse to create the same plot.
```{r}
p1 <- ggplot(data = DoseResponse_Data_Pig,
       aes(x = Dose, y = Activity)) +
  geom_point(col = "black", size = 0.5) + 
  geom_smooth(method = drm,col = "black", size = 0.5,linetype = "dashed", method.args = list(fct = L.4()), se = F) +
  scale_x_log10() +
  labs(title= "", x = "Dose (M)",  y = "Na+/K+-ATPase activity \n (% of non-inhibited control)") + 
  theme_classic()

p1 + coord_equal() #now we have square coordinate grid
p1 + theme(aspect.ratio = 1) #now the plot is square
p1 + coord_equal() + theme(aspect.ratio = 1)  
```

Exercise. Fit a log logistic dose-response curve to the monarch butterfly data with 4 parameters. What is the estimated effective IC50 dose? Plot the fitted curve both with base R and ggplot.

Solution.
```{r}
DoseResponse_Data_Pig<- read.csv("Figure3a.csv") %>% 
  dplyr::filter(Species_ATPase == "Monarch butterfly ATPase")

DoseResponse_Data_Monarch<- read.csv("Figure3a.csv") %>% 
  dplyr::filter(Species_ATPase == "Monarch butterfly ATPase")

head(DoseResponse_Data_Monarch)

DoseResponse_Data_Monarch.model <- drm(Activity ~ Dose, 
                                   data = DoseResponse_Data_Monarch, 
                                   fct = LL.4(fixed = c(NA, 0,100, NA)))

summary(DoseResponse_Data_Monarch.model)

ED(DoseResponse_Data_Monarch.model, c(50))

plot(DoseResponse_Data_Monarch.model,
     type = "all",
     lwd = 2,
     lty=3 ,
     col = "red",
     main="Monarch ATPase",
     xlab = "Dose (M)", 
     ylab = "Na+/K+-ATPase activity \n (% of non-inhibited control)")

p2 <- ggplot(data = DoseResponse_Data_Monarch,
       aes(x = Dose, y = Activity)) +
  geom_point(col = "red", size = 0.5) + 
  geom_smooth(method = drm,col = "red", size = 0.5,linetype = "dashed", method.args = list(fct = L.4()), se = F) +
  scale_x_log10() +
  labs(title= "", x = "Dose (M)",  y = "Na+/K+-ATPase activity \n (% of non-inhibited control)") + 
  theme_classic()

p2 + coord_equal() #now we have square coordinate grid
p2 + theme(aspect.ratio = 1) #now the plot is square
p2 + coord_equal() + theme(aspect.ratio = 1) 
```

